---
title: "Probability and Binomial Distribution"
author: "Barboza-Salerno"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Binomial Distribution

Recall that the formula for the binomia distribution is:

${n! \over k!(n-k)!} p^k(1-p)^{n-k}$

*Note:* This computes the probability of observing *exactly* x successes out of n.

### Example 1

Use the formula in R to calculate the following probability:

Two heads in three tosses of a fair coin.

```{r cars}
dbinom(x=2, size = 3, prob = .5)
```

Note, you can also calculate this 'by hand' using the factorial key word

```{r pressure, echo=FALSE}
n <- 3
k <- 2
p <- .5

factorial(n)/(factorial(k)*factorial(n - k)) * p^k * (1-p)^(n-k)
```

### You Try
Your turn. Suppose I want to form a research group of two students from three students. Suppose that a new therapy technique has a 33% success rate for depression. You are trying to figure out how many of your clients will be helped by this technique. You are a successful therapist with 100 clients. What is the probability that exactly 20 clients will be relieved of their depression? 

```{r eval = FALSE}
dbinom(x=20, size = 100, prob = 1/3)
```

What is the probability that *at least 20* will be relieved? For this, similar to before, we can use the pbinom function which calculates the probability of being less than or equal to a value.
```{r}
pbinom(20, size = 100, prob = 1/3)
```

and, the probability that 20 or more will be relieved...

```{r}
1 - pbinom(20, size = 100, prob = 1/3)
```

### Example 2
Suppose that the probability of a potato rotting after a month in the pantry is 1/3 (and that one potato rotting is independent of the others). Let’s say we have 5 potatoes in the pantry.

1) If we didn’t get around to eating our delicious potatoes for a month, what is the probability that all 5 are still good?

2) What is the probability that at least 3 potatoes are still good?

3) What is the probability that exactly two are still good?

```{r q1}
(dbinom(x=5, size = 5, prob = 1/3))
```

Alternatively, we can just multiple 1/3 five times.
```{r}
(1/3)^5
```


```{r q2}
dbinom(x=5, size = 5, prob = 1/3) + dbinom(x=4, size = 5, prob = 1/3) + dbinom(x=3, size = 5, prob = 1/3)
```

```{r q3}
dbinom(x=2, size = 5, prob = 1/3) 
```

### Example 3

The likelihood that a patient with a heart attack dies of the attack is 0.04 (i.e., 4 of 100 die of the attack). Suppose we have 5 patients who suffer a heart attack. 

1) What is the probability that all will survive?

2) What is the probability that no more than 1 person dies of a heart attack?

3) Carefully note all of the assumptions you made in using the binomial distribution to calculate these probabilities.

## Practice
Calculate the probability of each of the following:

1) Two heads in three tosses of a fair coin

2) Four heads in four tosses of a coin that is unfair, and the probability of a heads is 0.75

3) Three sixes in three rolls of a fair die

4) Less than 4 heads in 8 tosses of a fair coin (.3633)

5) More than 5 heads in 8 tosses of a fair coin (.1445)

5) (challege): How likely is it to pass (> .50) a multiple choice test if you have to guess the answer to each question, each question has four possible choices and the test has 10 total questions? Comment.

