---
title: "Z-scores"
author: "Barboza-Salerno"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Computing z-scores from p-values

To find the z-score associated with a p-value in R, we can use the `qnorm()` function, which uses the following syntax:
`qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)`

**Example**: Find the Z-score associated with the median of the distribution

```{r zmed}
qnorm(.5)  
```

**Example**: Find the Z-score for the p-value .6629

```{r z6}
qnorm(.6629)  
```

## Computing p-values from z-scores
To find the p-value associated with a z-score in R, we can use the `pnorm()` function, which uses the following syntax:
`pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)` 

What p-value is associated with a z-score = 0?

```{r}
pnorm(0)  
```

What p-value is associated with a z-score = .420?

```{r}
pnorm(.420)  
```

## Example
The length of a human pregnancy is normally distributed with a mean of 272 days with a standard deviation of 9 days (Bhat & Kushtagi, 2006).

1. State the random variable.
2. Find the probability of a pregnancy lasting more than 280 days.
3. Find the probability of a pregnancy lasting less than 250 days.
4. Find the probability that a pregnancy lasts between 265 and 280 days.
5. Find the length of pregnancy that 10% of all pregnancies last less than.
6. Suppose you meet a woman who says that she was pregnant for less than 250 days. Would this be unusual and what might you think?

#### Answers

1. The random variable, X, is the length of a human pregnancy.
2. We are looking for P(X > 280) which equals 1 - P(X <=280). Compute z-score and compute 1 - pnorm. (Also use common sense, does 89% seem right?)
3. We are looking for P(X<=250) which is easily computed
4. We are looking for P(265 < X < 280) (use common sense), and the compute P(X < 265) and then P(X < 280) and subtract the two

```{r q2}
# First compute z-score
(z_score <- (280 - 272)/9)
1-pnorm(z_score)
```


```{r q3}
# First compute z-score
(z_score <- (250 - 272)/9)
pnorm(z_score)
```

```{r q4}
# First compute z-score
(z_score_lower <- (265-272)/9)
(z_score_upper <- (280-272)/9)
pnorm(z_score_upper)-pnorm(z_score_lower)

```
```{r q5}
qnorm(.10)
# note we are now looking for X
# (X - 272)/9 = -1.282 --> re-arrange and solve for X
# X = 272 - 1.282(9)
272 - 1.282*9
```

